---
title: "Data Cleaning"
author: "Rebecca Folmer Schade"
date: "27 dec 2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# I'll set the working directory here as well
require("knitr")
opts_knit$set(root.dir = "~/Uni/Cultural_data_science/Exam/Open_data_project")

```

The raw data (imported from Survey Xact) is difficult to work with as the labels for columns and categories are not informative. This markdown is dedicated to remedying that.  

## Setup

To clean the data, I will be borrowing the powers of the R package *\'tidyverse\'*, which allows easier and more flexible data manipulation (though it is not used much). 

```{r Calling package, message=FALSE, warning=FALSE}
# calling the tidyverse package
library(tidyverse)
```

The raw data collected from the survey can be found within the *\'data\'* folder of this project.
Its file name is *\'dataset.csv\'*, and I load it into an object (of the class *\'tibble\'*) called *\'raw_data\'* below. 

```{r Loading data, message=FALSE, warning=FALSE}
# Loading raw data
raw_data <- read_csv("data/dataset.csv")
```

This leaves me with a data frame with 22 observations, 28 variables, and very uninformative labels.  

## Removing and Renaming Columns

Not all 28 variables are necessary for my analysis, so I will remove those I do not need for a better overview.  

```{r Colnames}
# Geting an overview of the columns of raw_data
colnames(raw_data)
```

Though these names are mostly informative, I know that *\'s_1\'* refers to a declaration of consent (which in unnecessary to include as it can be inferred by completing the survey), that *\'email\'* refers to the participants' emails (which is empty as they were not collected), and that columns *\'stato_1\'* to *\'stato_5\'* refers to the completion status of the surveys (of which I will only need to keep *\'stato_4\'* to exclude incomplete forms).  

Thus, I will be removing the following columns:   
*\'s_1\' - \'email\' - \'stato_1\' - \'stato_2\' - \'stato_3\' - \'stato_5\'*

```{r Removing Columns}
# Loading the raw_data into a new object to more easily make use of the original data later on
modified_data <- raw_data

# Making a vector of the names of the columns I wish to remove
to_remove <- c("s_1", "email", "stato_1", "stato_2", "stato_3", "stato_5")

# Subsetting out the columns which names match those in to_remove
modified_data <- modified_data[, !(names(modified_data) %in% to_remove)]
```

Now it is time to rename the remaining columns - See [the survey datapackage](https://github.com/RFSchade/CDS_exam_project/blob/main/Datapackage_survey.json) for proper documentation of the new variable names. 

```{r Renaming Columns}
# Renaming the remaining columns 
# > The remaining variables don't exactly have very informative names, so I'll change them
colnames(modified_data) <- c("age", 
                             "gender", 
                             "position", 
                             "years_experience", 
                             "field",
                             "familiarity",
                             "familiarity_scource",
                             "importance",
                             "definition_comment",
                             "share_after_pub",
                             "share_after_pub_comment",
                             "share_after_pub_freq",      
                             "share_pre_pub",
                             "share_pre_pub_comment",
                             "share_pre_pub_freq",
                             "public_data_freq",
                             "public_data_how",
                             "student_graded",
                             "teacher_graded",
                             "other_data_use",
                             "other_data_experience",
                             "status")

```

  

## Changing Labels Within the Data

Now that the names of the columns have become more informative, it becomes clear that labels of the categorical data within them are not. As such, I will change those as well. (for more through documentation of these labels, see: [the survey datapackage](https://github.com/RFSchade/CDS_exam_project/blob/main/Datapackage_survey.json))  

In the survey, the questions that specifies the categorical data has text options (though these categories in the current data are represented by numbers) - I will name the labels after those. To do this, I turn the variables from characters into a factors and label the levels. 

```{r Labeling Categories}
# Labeling 'gender'
modified_data$gender <- factor(modified_data$gender, 
                          levels = c(1:4), 
                          labels = c("male", 
                                     "female",
                                     "other", 
                                     "prefer not to say"))

# Labeling 'position'
modified_data$position <- factor(modified_data$position,
                            levels = c(1:4,6,7),
                            labels = c("academic",
                                       "archevist", 
                                       "Independent scholar",  
                                       "librarian",
                                       "phd researcher", 
                                       "student"))

# Labeling 'importance'
modified_data$importance <- factor(modified_data$importance,
                              levels = c(0:4),
                              labels = c("not important", 
                                         "slightly important", 
                                         "neutral", 
                                         "important", 
                                         "very important"))

# Labeling 'status'
modified_data$status <- factor(modified_data$status,
                          levels = c(0:1),
                          labels = c("incomplete", 
                                     "complete"))


# There are multiple questions where the answer key is yes/no, so I felt it would be prudent that I name them all in one go

# First, I make a list of the names of the columns with the yes/no answer key
binary <- c("familiarity", "share_after_pub", "share_pre_pub")
# then I make a for-loop to change all the labels in one go 
for (i in binary) {
  # I find that in these loops it is more safe to turn the variable into a facto prior to releveling. This    is not strictly necessary, but I get less errors when I do. 
  modified_data[[i]] <- as.factor(modified_data[[i]])
  modified_data[[i]] <- factor(modified_data[[i]], 
                               levels = c(2, 3), 
                               labels = c("yes",
                                          "no"))
}

# the questions for 'student_graded' and 'teacher_graded' also have the same answer key, so I will label these using the same method

# Making a list of column names 
grades <- c("student_graded", "teacher_graded")
# Making a for-loop 
for (i in grades) {
  modified_data[[i]] <- as.factor(modified_data[[i]])
  modified_data[[i]] <- factor(modified_data[[i]], 
                               levels = c(2:4), 
                               labels = c("yes", 
                                          "no", 
                                          "not relevant for me"))
}

# Several question about frequency also use the same answer key, so I will label these using a loop as well 
# Making a list of column-names
freq <- c("share_after_pub_freq", "share_pre_pub_freq", "public_data_freq", "other_data_use")
# Making a for-loop 
for (i in freq) {
  modified_data[[i]] <- as.factor(modified_data[[i]])
  modified_data[[i]] <- factor(modified_data[[i]], 
                               levels = c(1:5), 
                               labels = c("always", 
                                          "often", 
                                          "sometimes", 
                                          "rarely", 
                                          "never"))
}

```

  
## Work to do for later

I leave the cleaning at this for now - though there are more variables that could use some cleaning - such as *\'field\'* and *\'familiarity_scource\'* - this would be for the sake of standardization (as opposed to understandability). As time is short and I will not be visualizing these variables in this class project, I feel comfortable leaving the work to be done at a later date or for another project.  

## Saving the Dataset

I will save this dataset two times - once as it is (somewhat cleaner than the raw data, but still containing incomplete surveys) - and once when I have removed answers from participants who did not finish the survey. 

```{r}
# Saving the data as-is
write_csv(modified_data, "data/full_dataset.csv")

# Removing data from incomplete surveys using the filter() function from tidyverse
clean_data <- modified_data %>% filter(status == "complete")
# This leaves 18 rows 

# Removing the 'status' columns, as it is no longer needed
clean_data$status <- NULL

# Saving the clean dataset
write_csv(clean_data, "data/clean_dataset.csv")
```











