---
title: "Plotting"
author: "Rebecca Folmer Schade"
date: "27 dec 2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# I'll set the working directory here as well
require("knitr")
opts_knit$set(root.dir = "~/Uni/Cultural_data_science/Exam/Open_data_project")
```

The purpose of this project is to illustrate and shed further light on the intricacies of why open data is still lacking in scientific research, despite much time spendt as a hot topic in the debate on a chaning scintific process.  

To do this, I will attempt to plot data I have collected from a small survey of data sharing practices at Aarhus University, Denmark (though the data I have collected so far is not suficient for a statistical analysis). 

The goal is for these plots to be:  
- easily readable  
- informative  
- comparable to each-other

## Setup

Though the data has been cleaned (see [Data_cleaning_markdown.Rmd](https://github.com/RFSchade/CDS_exam_project/blob/main/Data_cleaning_markdown.Rmd)), I will need help from the *\'tidiverse\'* package to manipulate it into a shape more beneficient to visialisation, as well as plotting the data. 

```{r Calling package, message=FALSE, warning=FALSE}
# calling the tidyverse package
library(tidyverse)
```

The cleaned dataset can be found within the *\'data\'* folder of this project.
Its file name is *\'clean_dataset.csv\'*, and I load it into an object (of the class *\'tibble\'*) called *\'aarhus_survey\'* below.

```{r Loading data, message=FALSE, warning=FALSE}
# Loading clean dataset
aarhus_survey <- read_csv("data/clean_dataset.csv")
```


## Defining a theme and color scheme

First I want to define some commonalieties between the plots. I intend to do this using a custom *\'ggplot\'* theme and colorscheme. 

```{r Theme, echo=FALSE}
# Creating a GGplot theme using 'theme_minimal' as a basis
showcase_theme <- theme_minimal() + 
  theme(panel.grid.minor = element_blank()) + # Removing the minor background grid
  theme(panel.grid.major.y = element_blank()) + # Removing part of the major background grid
  theme(panel.grid.major.x = element_line(colour = "#989898")) + # changing the color of the remaining grid
  theme(aspect.ratio = 1/5) + # Changing the aspect ratio between the x and y axis
  theme(legend.position = "top") + # Moving the legend to the top of the plot 
  theme(legend.title = element_blank()) + # Removing the legend title
  theme(legend.text = element_text(size = 8)) + # Changing the size of the legend text
  theme(legend.key.width = unit(2, "cm")) + # standardizing the length of the legend keys
  theme(plot.title = element_text(hjust = 0.5)) # adjusting the position of the tilt of the plot

# Defining a list of colors for likert scale plots
likert_palette <- c("#0965C5", "#99CAF3", "#BFCBD5", "#F5B09E", "#B13311")

# Defining a list of colors for plots with a binary answer key 
binary_palette <- c("#0965C5", "#B13311")
```

## Manipulating data

The data in *\'aarhus_survey\'*, while clean, is not in the right format for plotting. In this section, I will change that. 

```{r Data wrangling}
# Creating a new column for separating the data into two groups - Students and Academics
aarhus_survey <- aarhus_survey %>% 
  mutate(demographic = ifelse(position == "student", "Students", "Academics"))

# Subsetting the dataframe to the questions I need for visualization
sub_survey <- aarhus_survey[,c(8,10,12,13,15,16,20,22)]

# Making a list of the new names of the columns
# Some of these names only contain the last part of the question
# That is because the first part will be in the plot title
colnames(sub_survey) <- c("How important is open data?", 
                          "Share data after publication?", # Starts "is it important to" 
                          "Share your data after publication?", # Starts "How often do you"
                          "Share data before publication?", # Starts "is it important to"
                          "Share your data before publication?", # Starts "How often do you"
                          "Make your data public?", # Starts "How often do you"
                          "Use other people\'s data?", # Starts "How often do you"
                          "Demographic")

```

I will be making three different plots; One with importance rated in a likert scale, one with importance rated with a binary scale, and one with frequency rated in a likert scale. 

```{r}
#=====> Making data from the importance question the right format for plotting
# Subsetting only the columns I need
importance <- sub_survey[,c(1,8)]

# Making the data Long format
importance_long <- importance %>% gather(question, category, 1, factor_key=TRUE)

# Summarising the data by question, category and demographic 
importance_agg <- importance_long %>% 
  group_by(question, category, Demographic) %>% 
  dplyr::summarise(total = n())

# Adding a column showing percentages for plotting
importance_sum <- importance_agg %>% 
  group_by(Demographic) %>% # Grouping by demographic...
  mutate(countT = sum(total)) %>% # ...so the total number of answers are calculated by demographic
  group_by(category, add=TRUE) %>% # Grouping by category...
  mutate(per=round(100*total/countT,2)) # ..so the percenteges are calculated by category and demographic

# Turning category into a factor so I can order the labels after importance for plotting 
importance_sum$category <- as.factor(importance_sum$category)
# Ordering the labels
importance_sum$category <- 
  ordered(importance_sum$category, 
          levels = c("not important", 
                     "slightly important",
                     "neutral", 
                     "important", 
                     "very important"))


#=====> Making data from the binary questions the right format for plotting
# Subsetting only the columns I need
binary <- sub_survey[,c(2,4,8)]

# Making the data Long format
binary_long <- binary %>% gather(question, category, 1:2, factor_key=TRUE)

# Summarising the data by question, category and demographic 
binary_agg <- binary_long %>% group_by(question, category, Demographic) %>% dplyr::summarise(total = n())

# Adding a column showing percentages for plotting
binary_sum <- binary_agg %>%
  group_by(question, Demographic) %>% # Grouping by demographic and question...
  mutate(countT= sum(total)) %>% # ...so the total number of answers are calculated by the two variables
  group_by(category, add=TRUE) %>% # Grouping by category...
  mutate(per=round(100*total/countT,2)) # ..so the percenteges are calculated by all three varaibles

# Turning category into a factor so I can order the labels for plotting 
binary_sum$category <- as.factor(binary_sum$category)
# Ordering the labels
binary_sum$category <- 
  ordered(binary_sum$category, 
          levels = c("no", "yes"))


#=====> Making the data from the frequency questions the right format for plotting
# Subsetting only the columns I need and filtering the students out of the sample
freq <- sub_survey[,c(3,5,6,7,8)] %>% filter(Demographic == "Academics")

# Making the data Long format
freq_long <- freq %>% gather(question, category, 1:4, factor_key=TRUE)

# Summarising the data by question and category
freq_agg <- freq_long %>% group_by(question, category) %>% dplyr::summarise(total = n())

# Adding a column showing percentages for plotting
freq_sum <- freq_agg %>%
  group_by(question) %>% # Grouping by question...
  mutate(countT= sum(total)) %>% # ...so the total number of answers are calculated by question
  group_by(category, add=TRUE) %>% # Grouping by category...
  mutate(per=round(100*total/countT,2)) # ..so the percenteges are calculated by question and category

# Turning category into a factor so I can order the labels after frequency for plotting
freq_sum$category <- as.factor(freq_sum$category)
# Ordering the labels
freq_sum$category <- 
  ordered(freq_sum$category, 
          levels = c("never", "rarely", "sometimes", "often", "always"))

```

## Plotting 

For importance 

```{r, message=FALSE}
# Creating the plot
ggplot(data = importance_sum, aes(x =question , y = per, fill = category)) + 
  # Making it a bar plot
  geom_bar(stat="identity") +
  # Applying custom color scheme
  scale_fill_manual (values=likert_palette, 
                     labels=levels(importance_sum$category), 
                     drop = FALSE, 
                     na.value = "#4F4F4F") +
  # Flipping the bars from vertical to horizontal 
  coord_flip() + 
  # Removing axis titles for simplicity and adding a plot title
  labs(x = element_blank(), y = element_blank(), title = "How important is open data?") + 
  # Adding the custom theme
  showcase_theme + 
  # Removing the y axis ticks to save space
  theme(axis.text.y = element_blank()) +
  # Adding %-signs to the x axis
  scale_y_continuous(labels = function(x) paste0(x, "%")) + 
  # Changing the position of the legend 
  guides(fill = guide_legend(nrow = 1, label.position = "top", reverse = TRUE)) + 
  # Facetting the plot by demographic
  facet_wrap(~Demographic) 

# Saving the plot
ggsave("output/importance_plot.jpg")
```

For binary data 

```{r, message=FALSE}
# Creating the plot
ggplot(data = binary_sum, aes(x =question , y = per, fill = category)) + 
  # Making it a bar plot
  geom_bar(stat="identity") + 
  # Applying custom color scheme
  scale_fill_manual (values=binary_palette, 
                     labels=levels(binary_sum$category), 
                     drop = FALSE, 
                     na.value = "#4F4F4F") +
  # Flipping the bars from vertical to horizontal
  coord_flip() +
  # Removing axis titles for simplicity and adding a plot title
  labs(x = element_blank(), y = element_blank(), title = "Is it important to...") +
  # Adding the custom theme
  showcase_theme + 
  # Adding %-signs to the x axis
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  # Changing the position of the legend 
  guides(fill = guide_legend(nrow = 1, label.position = "top", reverse = TRUE)) +
  # Facetting the plot by demographic, and specifying the direction of the facets
  facet_wrap(~Demographic, dir = "v")

# Saving the plot
ggsave("output/binary_plot.jpg")
```

For frequency data

```{r, message=FALSE}
# Creating the plot
ggplot(data = freq_sum, aes(x =question , y = per, fill = category)) + 
  # Making it a bar plot
  geom_bar(stat="identity") +
  # Applying custom color scheme
  scale_fill_manual (values=likert_palette, 
                     labels=levels(freq_sum$category), 
                     drop = FALSE, 
                     na.value = "#4F4F4F") +
  # Flipping the bars from vertical to horizontal
  coord_flip() + 
  # Removing axis titles for simplicity and adding a plot title
  labs(x = element_blank(), y = element_blank(), title = "How often do you...") +
  # Adding the custom theme
  showcase_theme + 
  # Changing the plot's aspect ratio betweewn x and y axis and adjusting the size of the legend 
  theme(aspect.ratio = 2/4, legend.key.width = unit(1.5, "cm")) + 
  # Adding %-signs to the x axis
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  # Changing the position of the legend 
  guides(fill = guide_legend(nrow = 1, label.position = "top", reverse = TRUE)) 

# Saving the plot
ggsave("output/frequency_plot.jpg")
  
```

## Descriptive stats

To describe my sample in the report, I will need a table of descriptive stats. This section is provided to create it, and to fish out whatever other descriptive stats are relevant. 

Before presenting readers to the table, I want to rattle off the stats before I direct there reader to the table.

```{r, message=FALSE}
# Calculating the mean age of participants, rounded to 2 decimals
round(mean(aarhus_survey$age), digits = 2)
# > 26.94
# Calculating the standard deviation of that mean, rounded to 2 decimals
round(sd(aarhus_survey$age), digits = 2)
# > 6.38

# Summaraisng the dataframe to get the gender distribution
gender <- aarhus_survey %>% group_by(gender) %>% summarise(number = n())
# > 10 female, 7 male, 1 other
```

I want to make a table of the distribution the academic fields of the participants, but to do so I first need to group them, as these answers are in free text. 

```{r, message=FALSE}
# Looking at the nr. of unique answers to the question about participants' academic field
unique(aarhus_survey$field)
# > There are 10 unique answers, however, several of them are different spelling or versions of the same field. This will have to be changed. 

# Making all answes lower case to remove variance between answers 
aarhus_survey$field <- tolower(aarhus_survey$field)
# Checking if that made my job easier 
unique(aarhus_survey$field)
# It did. Now there are only 8. 
# Mashing "psychology" and "psychology and neuroscience" into one category
aarhus_survey$field <- str_replace_all(aarhus_survey$field, 
                                       pattern = "psychology$", 
                                       replacement = "psychology and neuroscience")
# Checking if that worked
unique(aarhus_survey$field)
# It did - now there are only 7
# As I do not know what the answer "c" could refer to, I am changing it to NA
is.na(aarhus_survey$field) <- aarhus_survey$field == "c"

```

I will be creating a table of the distribution of academic fileds using the edited list in this next section: 

```{r}
# Summarising the data according to field 
field <- aarhus_survey %>% group_by(field) %>% summarise(number = n())
# Printing the result
print(field)
```
lastly, I want to create a table showing some descriptive stats according to the participants' position.

```{r}
# Creating summary dataframe 
position <- aarhus_survey %>% group_by(position) %>% # grouping according to position
  summarise(number = n(), # Getting number of answers  
              mean_exprerience = round(mean(years_experience),2), # Getting the mean years of expreience
              sd_expreience = round(sd(years_experience),2)) %>% # Getting the standard deviation (sd)
  mutate(mean_exp_total = round(mean(mean_exprerience),2), # mutating a column for the total mean experience
         sd_exp_total = round(sd(mean_exprerience),2)) # mutating a column for the sd of the total mean
# Printing result
print(position)

```


